Learning From Massive Noisy Labeled Data for Image Classification

### 2015-04-09 paper.tex
Polish the whole paper

\renewcommand\Authands{ and }

@@

Deep learning with massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2013panda}, scene classification~\cite{farabet2013learning}, \etc. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people first initialize the parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2013learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a Convolutional Neural Network (CNN) from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

@@

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the ground truths and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modifies a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumes noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit the real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. It shows that images tend to be mislabeled may share similar visual patterns. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for many similar clothes. Based on these observations, we introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

The proposed probabilistic model build the relations among the image, noisy label, ground truth label, and the noise type, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels.

Our contribution comes from three aspects. First, we study the cause of noisy labels in real-world data and describe it with a novel probabilistic model. Second, we integrate the model into existing deep learning framework to help detect and correct wrong labels. We study different training strategies to make the CNNs learn from better supervisions. Finally, we collect a large-scale clothing dataset with both noisy and clean labels, which will be released for academic use.

@@

{\bf Semi-supervised learning:} Apart from directly learning with label noise, some semi-supervised learning algorithms are developed to utilize weakly labeled or even unlabeled data. Label Propagation method~\cite{zhu2002learning} explicitly uses ground truths of well-labeled data to classify unlabeled samples. However, it suffers from computing pairwise distance, which has quadratic complexity with the number of samples and cannot be applied on large-scale datasets. Chen \etal~\cite{chen2013fast} tackled this problem in the scenario of image tagging by a co-regularized duo classifier that can efficiently use both the image and incomplete tags. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they need extra information about whether a pair of unlabeled images belong to the same class, which cannot be obtained in our problem.

{\bf Transfer learning:} The success of Convolutional Neural Networks (CNNs) lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing the CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning it on the smaller dataset of specific task~\cite{krizhevsky2012imagenet, oquab2013learning, azizpour2014generic,donahue2013decaf}. Nevertheless, this transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothing classification, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it from an ImageNet pretrained model only on the clean labels.

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}.~\cite{mnih2012learning} builds a simple noise model for aerial images but only considers binary classification.~\cite{larsen1998design} assumes label noise is independent from the true class label which is a simple and a specific case.~\cite{sukhbaatar2014learning} generalizes from them by considering multi-class classification and modeling class dependent noise, but they assume the noise is conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels. Our work can be viewed as a generalization of~\cite{sukhbaatar2014learning,natarajan2013learning} and our model is flexible enough to not only class dependent noise but also input dependent noise.
% section related_work (end)

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on the problem of learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_n=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Figure~\ref{fig:noise_pgm}.

@@

Despite the observed image $\vec{x}$ and the noisy label $\tilde{\vec{y}}$, we exploit two discrete latent variables --- $\vec{y}$ and $\vec{z}$ --- to represent the ground truth and the type of label noise, respectively. Both $\tilde{\vec{y}}$ and $\vec{y}$ are $L$-dimensional binary random variables in 1-of-$L$ fashion, \ie, only one element is equal to 1 while others are all 0.

On the other hand, the latent variable $\vec{z}$ representing label noise type is an 1-of-3 binary random variable. It is associated with three different semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$
\end{enumerate}

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \mat{C}\vec{y} & \text{if } \vec{z}_3 = 1
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are one), and $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denoting the confusion probability between class $i$ and $j$. Then we can derive from Figure~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$:
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x})
\end{equation}

@@

\subsection{Learning the Parameters} % (fold)
\label{sub:learning_the_parameters}
We exploit two CNNs to model $p(\vec{y} | \vec{x})$ and $p(\vec{z} | \vec{x})$ separately. Denote the parameter set of each deep model by $\theta_1$ and $\theta_2$. Our goal is to find the optimal $\theta = \theta_1 \cup \theta_2$ that maximize the incomplete log-likelihood $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$. The Expectation-Maximization algorithm is used to solve this problem iteratively.

For any probability distribution $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})$, we can derive a lower bound of the incomplete log-likelihood by
\begin{equation}
\begin{aligned}
    \log p(\tilde{\vec{y}} | \vec{x}; \theta) &= \log \sum_{\vec{y},\vec{z}} p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    &\geq \sum_{\vec{y},\vec{z}} q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) \log \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)}{q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})}
\end{aligned}
\end{equation}
\textbf{E-Step} The difference between $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$ and its lower bound is the Kullback-Leibler divergence $\mathrm{KL}\left(q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) || p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)\right)$, which is equal to zero if and only if $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) = p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)$. Therefore, in each iteration $t$, we first compute the posterior of latent variables using current parameters $\theta^{(t)}$:
\begin{multline} \label{eq:posterior}
    p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) = \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta^{(t)})}{p(\tilde{\vec{y}} | \vec{x}; \theta^{(t)})} \\
    = \frac{p(\tilde{\vec{y}}|\vec{y},\vec{z};\theta^{(t)}) p(\vec{y}|\vec{x};\theta^{(t)}) p(\vec{z}|\vec{x};\theta^{(t)})}{\sum_{\vec{y}',\vec{z}'} p(\tilde{\vec{y}}|\vec{y}',\vec{z}';\theta^{(t)}) p(\vec{y}'|\vec{x};\theta^{(t)}) p(\vec{z}'|\vec{x};\theta^{(t)})}
\end{multline}
Then the expected complete log-likelihood can be written as
\begin{equation}
    Q(\theta; \theta^{(t)}) = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)
\end{equation}
\textbf{M-Step} Since deep neural networks are exploited to model the probability $p(\vec{y}|\vec{x};\theta_1)$ and $p(\vec{z}|\vec{x};\theta_2)$, we perform gradient ascent on $Q$:
\begin{multline} \label{eq:gradient}
    \frac{\partial Q}{\partial \theta} = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta} \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    = \sum_{\vec{y}} p(\vec{y} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_1} \log p(\vec{y}|\vec{x};\theta_1) + \\
      \sum_{\vec{z}} p(\vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_2} \log p(\vec{z}|\vec{x};\theta_2)
\end{multline}

@@

\subsection{Estimating the Matrix $\mat{C}$} % (fold)
\label{sub:confusion_estimation}
Notice that we do not set parameters to the conditional probability $p(\tilde{\vec{y}}|\vec{y},\vec{z})$ in Eq~\eqref{eq:likelihood} and keep it unchanged during the learning process. Because without other regularizations, learning all the three parts could lead to trivial solutions. For example, the network will always predict $\vec{y}_1 = 1$, $\vec{z}_3 = 1$, and the matrix $\mat{C}$ is learned to make $\mat{C}_{1i} = 1$ for all $i$. To avoid this degeneration, we choose to give $\mat{C}$ a good estimation while let the deep model learn to predict $p(\vec{y} | \vec{x})$ and $p(\vec{z} | \vec{x})$ for each input image.

We estimate $\mat{C}$ on a relatively small dataset $\mathcal{D}_c = \{(\vec{x},\vec{y},\tilde{\vec{y}})\}_N$, where we have $N$ images with both clean and noisy labels. As prior information about $\vec{z}$ is not available, we just solve the following optimization problem:
\begin{equation} \label{eq:opt_all}
    \max_{\mat{C},\vec{z}^{(1)},\cdots,\vec{z}^{(N)}} \sum_{i=1}^N \log p(\tilde{\vec{y}}^{(i)}|\vec{y}^{(i)},\vec{z}^{(i)})
\end{equation}
Obviously, sample $i$ contributes nothing to the optimal $\mat{C}^*$ if $\vec{y}^{(i)}$ and $\tilde{\vec{y}}^{(i)}$ are equal. So that we ignore those samples and reinterpret the problem in another form by exploiting Eq~\ref{eq:likelihood}:
\begin{equation}
\begin{aligned}
\max_{\mat{C},\vec{t}} \quad & E=\sum_{i=1}^{N'} \log \alpha^{\vec{t}_i} + \log (\tilde{\vec{y}}^{(i)T} \mat{C} \vec{y}^{(i)})^{1-\vec{t}_i}\\
\text{subject to} \quad & \mat{C} \text{ is a stochastic matrix of size } L\times L \\
                    & \vec{t} \in \{0,1\}^{N'}
\end{aligned}
\end{equation}
where $\alpha=\frac{1}{L-1}$ and $N'$ is the number of remaining samples. The semantic meaning of the above formulation is that we need to assign each $(\vec{y},\tilde{\vec{y}})$ pair the optimal noise type, while finding the optimal $\mat{C}$ simultaneously.

>>>>

\renewcommand\Authands{, and }

@@

Deep learning with a massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2013panda}, scene classification~\cite{farabet2013learning}, \etc. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people first initialize the parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2013learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

@@

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modifies a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumes noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit the real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. It shows that images tend to be mislabeled may share similar visual patterns. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for many similar clothes. Based on these observations, we introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

The proposed probabilistic model builds the relations among the image, noisy label, ground truth label, and the noise type, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels.

Our contribution comes from three aspects. First, we study the cause of noisy labels in real-world data and describe it with a novel probabilistic model. Second, we integrate the model into a deep learning framework to help detect and correct wrong labels. We study different training strategies to make the CNNs learn from better supervisions. Finally, we collect a large-scale clothing dataset with both noisy and clean labels, which will be released for academic use.

@@

{\bf Semi-supervised learning:} Apart from directly learning with label noise, some semi-supervised learning algorithms are developed to utilize weakly labeled or even unlabeled data. Label Propagation method~\cite{zhu2002learning} explicitly uses ground truths of well labeled data to classify unlabeled samples. However, it suffers from computing pairwise distance, which has quadratic complexity with the number of samples thus cannot be applied on large-scale datasets. Chen \etal~\cite{chen2013fast} tackled this problem in the scenario of image tagging by a co-regularized duo classifier that can efficiently use both the image and incomplete tags. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they need extra information about whether a pair of unlabeled images belong to the same class, which cannot be obtained in our problem.

{\bf Transfer learning:} The success of CNNs lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning it on the smaller dataset of specific task~\cite{krizhevsky2012imagenet, oquab2013learning, azizpour2014generic,donahue2013decaf}. Nevertheless, this transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothing classification, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels.

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}.~\cite{mnih2012learning} builds a simple noise model for aerial images but only considers binary classification.~\cite{larsen1998design} assumes label noise is independent from the true class label which is a simple and special case.~\cite{sukhbaatar2014learning} generalizes from them by considering multi-class classification and modeling class dependent noise, but they assume the noise is conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels. Our work can be viewed as a generalization of~\cite{sukhbaatar2014learning,natarajan2013learning} and our model is flexible enough to not only class dependent noise but also input dependent noise.
% section related_work (end)

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_n=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with the $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Figure~\ref{fig:noise_pgm}.

@@

Despite the observed image $\vec{x}$ and the noisy label $\tilde{\vec{y}}$, we exploit two discrete latent variables --- $\vec{y}$ and $\vec{z}$ --- to represent the true label and the label noise type, respectively. Both $\tilde{\vec{y}}$ and $\vec{y}$ are $L$-dimensional binary random variables in 1-of-$L$ fashion, \ie, only one element is equal to 1 while others are all 0.

On the other hand, the label noise type $\vec{z}$ is an 1-of-3 binary random variable. It is associated with three different semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$
\end{enumerate}

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \tilde{\vec{y}}^{T}\mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}\tilde{\vec{y}}^{T}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \tilde{\vec{y}}^{T}\mat{C}\vec{y} & \text{if } \vec{z}_3 = 1
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are one), and $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denoting the confusion probability between class $i$ and $j$. Then we can derive from Figure~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$:
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x})
\end{equation}

@@

\subsection{Learning the Parameters} % (fold)
\label{sub:learning_the_parameters}
We exploit two CNNs to model $p(\vec{y} | \vec{x})$ and $p(\vec{z} | \vec{x})$ separately. Denote the parameter set of each deep model by $\theta_1$ and $\theta_2$. Our goal is to find the optimal $\theta = \theta_1 \cup \theta_2$ that maximize the incomplete log-likelihood $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$. The Expectation-Maximization algorithm is used to iteratively solve this problem.

For any probability distribution $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})$, we can derive a lower bound of the incomplete log-likelihood
\begin{equation}
\begin{aligned}
    \log p(\tilde{\vec{y}} | \vec{x}; \theta) &= \log \sum_{\vec{y},\vec{z}} p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    &\geq \sum_{\vec{y},\vec{z}} q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) \log \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)}{q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})}
\end{aligned}
\end{equation}
\textbf{E-Step} The difference between $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$ and its lower bound is the Kullback-Leibler divergence $\mathrm{KL}\left(q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) || p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)\right)$, which is equal to zero if and only if $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) = p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)$. Therefore, in each iteration $t$, we first compute the posterior of latent variables using current parameters $\theta^{(t)}$:
\begin{multline} \label{eq:posterior}
    p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) = \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta^{(t)})}{p(\tilde{\vec{y}} | \vec{x}; \theta^{(t)})} \\
    = \frac{p(\tilde{\vec{y}}|\vec{y},\vec{z};\theta^{(t)}) p(\vec{y}|\vec{x};\theta^{(t)}) p(\vec{z}|\vec{x};\theta^{(t)})}{\sum_{\vec{y}',\vec{z}'} p(\tilde{\vec{y}}|\vec{y}',\vec{z}';\theta^{(t)}) p(\vec{y}'|\vec{x};\theta^{(t)}) p(\vec{z}'|\vec{x};\theta^{(t)})}
\end{multline}
Then the expected complete log-likelihood can be written as
\begin{equation}
    Q(\theta; \theta^{(t)}) = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)
\end{equation}
\textbf{M-Step} We exploit two CNNs to model the probability $p(\vec{y}|\vec{x};\theta_1)$ and $p(\vec{z}|\vec{x};\theta_2)$, respectively. Thus the gradient of $Q$ \wrt $\theta$ can be decoupled into two parts:
\begin{multline} \label{eq:gradient}
    \frac{\partial Q}{\partial \theta} = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta} \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    = \sum_{\vec{y}} p(\vec{y} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_1} \log p(\vec{y}|\vec{x};\theta_1) + \\
      \sum_{\vec{z}} p(\vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_2} \log p(\vec{z}|\vec{x};\theta_2)
\end{multline}

@@

\subsection{Estimating the Matrix $\mat{C}$} % (fold)
\label{sub:confusion_estimation}
Notice that we do not set parameters to the conditional probability $p(\tilde{\vec{y}}|\vec{y},\vec{z})$ in Eq~\eqref{eq:likelihood} and keep it unchanged during the learning process. Because without other regularizations, learning all the three parts in Eq~\eqref{eq:joint_distribution} could lead to trivial solutions. For example, the network will always predict $\vec{y}_1 = 1$, $\vec{z}_3 = 1$, and the matrix $\mat{C}$ is learned to make $\mat{C}_{1i} = 1$ for all $i$. To avoid this degeneration, we estimate $\mat{C}$ on a relatively small dataset $\mathcal{D}_c = \{(\vec{x},\vec{y},\tilde{\vec{y}})\}_N$, where we have $N$ images with both clean and noisy labels. As prior information about $\vec{z}$ is not available, we just solve the following optimization problem:
\begin{equation} \label{eq:opt_all}
    \max_{\mat{C},\vec{z}^{(1)},\cdots,\vec{z}^{(N)}} \sum_{i=1}^N \log p(\tilde{\vec{y}}^{(i)}|\vec{y}^{(i)},\vec{z}^{(i)})
\end{equation}
Obviously, sample $i$ contributes nothing to the optimal $\mat{C}^*$ if $\vec{y}^{(i)}$ and $\tilde{\vec{y}}^{(i)}$ are equal. So that we discard those samples and reinterpret the problem in another form by exploiting Eq~\ref{eq:likelihood}:
\begin{equation}
\begin{aligned}
\max_{\mat{C},\vec{t}} \quad & E=\sum_{i=1}^{N'} \log \alpha^{\vec{t}_i} + \log (\tilde{\vec{y}}^{(i)T} \mat{C} \vec{y}^{(i)})^{1-\vec{t}_i}\\
\text{subject to} \quad & \mat{C} \text{ is a stochastic matrix of size } L\times L \\
                    & \vec{t} \in \{0,1\}^{N'}
\end{aligned}
\end{equation}
where $\alpha=\frac{1}{L-1}$ and $N'$ is the number of remaining samples. The semantic meaning of the above formulation is that we need to assign each $(\vec{y},\tilde{\vec{y}})$ pair the optimal noise type, while finding the optimal $\mat{C}$ simultaneously.

Next, we will show that the problem can be solved by a simple yet efficient algorithm in $O(N'+L^2)$ time complexity. Denote the optimal solution by $\mat{C}^*$ and $\vec{t}^*$
\begin{theorem}
$\mat{C}^*_{ij} \neq 0 \Rightarrow \mat{C}^*_{ij} > \alpha, \forall i,j \in \{1,\dots,L\}$
\end{theorem}
\begin{proof}
Suppose there exists some $i,j$ such that $0 < \mat{C}^*_{ij} \leq \alpha$. Then we conduct following operations. First, we set $\mat{C}^*_{ij}=0$ while adding its original value to other elements in column $j$. Second, for all the samples $n$ where $\tilde{\vec{y}}^{(n)}_i=1$ and $\vec{y}^{(n)}_j=1$, we set $\vec{t}_n$ to 1. The resulting $E$ will get increased, which leads to a contradiction.
\end{proof}
\begin{theorem}
$(\tilde{\vec{y}}^{(i)},\vec{y}^{(i)})=(\tilde{\vec{y}}^{(j)},\vec{y}^{(j)}) \Rightarrow \vec{t}^*_i=\vec{t}^*_j, \forall i,j \in \{1,\dots,N'\}$
\end{theorem}
\begin{proof}
Suppose $\tilde{\vec{y}}^{(i)}_k=\tilde{\vec{y}}^{(j)}_k=1$ and $\vec{y}^{(i)}_l=\vec{y}^{(j)}_l=1$ but $\vec{t}^*_i \neq \vec{t}^*_j$. From Theorem 1 we know that elements in $\mat{C}^*$ is either 0 or greater than $\alpha$. If $\mat{C}^*_{kl}=0$, we can set $\vec{t}^*_i=\vec{t}^*_j=1$, otherwise we can set $\vec{t}^*_i=\vec{t}^*_j=0$. In either case the resulting $E$ will get increased, which leads to a contradiction.
\end{proof}
\begin{theorem}
$\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} > \alpha \Leftrightarrow \vec{t}^*_i=0$ and $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} = 0 \Leftrightarrow \vec{t}^*_i=1, \forall i \in \{1,\dots,N'\}$
\end{theorem}
\begin{proof}
The first part is straight forward. For the second part, $\vec{t}^*_i=1$ implies $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} \leq \alpha$. By using Theorem 1 we know that $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} = 0$.
\end{proof}

Notice that if the true label is class $i$ while the noisy label is class $j$, then it can only affect the value of $\mat{C}_{ij}$. Thus each column of $\mat{C}$ can be optimized separately. Theorem 1 further shows that samples with same pair of $(\tilde{\vec{y}},\vec{y})$ share a same noise type. Then what really matters is the frequencies of all the $L\times L$ pairs of $(\tilde{\vec{y}},\vec{y})$. Considering a particular column $\vec{c}$, suppose there are $M$ samples affecting this column. We can count the frequencies of noisy label class 1 to $L$ as $m_1,\dots,m_L$ and might as well assume $m_1\geq m_2\geq \cdots \geq m_L$. The problem is then converted to
\begin{equation}
\begin{aligned}
\max_{\vec{c},\vec{t}} \quad & E=\sum_{k=1}^{L} m_k \left(\log \alpha^{\vec{t}_k} + \log \mat{c}_k^{1-\vec{t}_k}\right)\\
\text{subject to} \quad & \vec{c} \in [0,1]^L, \sum_{k=1}^{L} \vec{c}_k = 1 \\
                    & \vec{t} \in \{0,1\}^{L}
\end{aligned}
\end{equation}

Due to the rearrangement inequality, we can prove that in the optimal solution,
\begin{equation}
    \max(\alpha,\vec{c}^*_1) \geq \max(\alpha,\vec{c}^*_2) \geq \cdots \geq \max(\alpha,\vec{c}^*_L)
\end{equation}
Then by using Theorem 3, there must exist a $k^* \in \{1,\dots,L\}$ such that
\begin{equation} \label{eq:opt_t}
\begin{aligned}
    &\vec{t}^*_i = 0, i = 1,\dots,k^* \\
    &\vec{t}^*_i = 1, i = k^*+1,\dots,L
\end{aligned}
\end{equation}
This also implies that only the first $k^*$ elements of $\vec{c}^*$ have nonzero values (greater than $\alpha$ actually). Furthermore, if $k^*$ is known, finding the optimal $\vec{c}^*$ is to solve the following problem:
\begin{equation}
\begin{aligned}
\max_{\vec{c}} \quad & E=\sum_{k=1}^{k^*} m_k \log \mat{c}_k\\
\text{subject to} \quad & \vec{c} \in [0,1]^L, \sum_{k=1}^{k^*} \vec{c}_k = 1
\end{aligned}
\end{equation}
whose solution is
\begin{equation} \label{eq:opt_c}
\begin{aligned}
    &\vec{c}^*_i = \frac{m_i}{\sum_{k=1}^{k^*}{m_k}}, i = 1,\dots,k^* \\
    &\vec{c}^*_i = 0, i = k^*+1,\dots,L
\end{aligned}
\end{equation}
